
---
title: "Clustering Method for NCI60 data set."
author: "By Puspanjali Subudhi"
output:
  html_document:
    highlight: tango
    theme: lumen
  pdf_document: default
  word_document: default
---
---
output:
  html_document: default
  word_document: default
  pdf_document: default
---


***OBJECTIVE:***


***Using the clustering method to "NCI60"(cancer cell line microarray) data, with the goal of finding out whether or not the observations cluster into the distinct types of cancer. A comparison between K_means and hierarchical clustering.***



```{r}
#NCI60 data set info
library(ISLR)
nci.labs=NCI60$labs
nci.data=NCI60$data
dim(nci.data)
```

```{r}
nci.labs[1:4]
table(nci.labs)
nci.labs
```

```{r}
sd.data=scale(nci.data) # standardized the variables (mean zero and standard deviation one).
```
***Comparison between complete, average and single linkage.***

```{r}
par(mfrow=c(1,1))
data.dist = dist(sd.data)
plot(hclust(data.dist),labels=nci.labs,main="Complete Linkage",
xlab="",sub="",ylab = "",cex=".5",col="blue") #Hierarchical clustering of the observations using Complete Linkage.

```


```{r}
par(mfrow=c(1,1)) 
plot(hclust(data.dist,method="average"),labels=nci.labs,main="AverageLinkage",
xlab="",sub="",ylab = "",cex=".5",col="blue") #Hierarchical clustering of the observations using Average linkage.
```

```{r}
par(mfrow=c(1,1))
plot(hclust(data.dist,method="single"),labels=nci.labs,main="SingleLinkage",
xlab="",sub="",ylab= "",cex=".5",col="blue") #Hierarchical clustering of the observations using Single Linkage.
```




***OBSERVATIONS:***

***Single clusters tend to yield trailing clusters: very large clusters into which individual observations attached one by one.***

***On the other hand, complete and average linkage tends to yield more balanced and attractive clusters.***

***For this reason complete and average linkage are preferred to single linkage. Cell lines within a single cancer type do tend to cluster together, although the clustering is not perfect.***


```{r}
hc.out=hclust(dist(sd.data))
hc.clusters=cutree(hc.out,4)
table(hc.clusters,nci.labs)
```

***We can see a clear pattern that all the leukemia cell lines fall under cluster 3 where the breast cancer cells are spread out over three different cluster.***


```{r}
plot(hc.out,labels=nci.labs,main="Cluster Dendrogram",xlab="",sub="",ylab = "",cex=".5", col="25",cex=".5")
abline(h=139, col="red")
```

***The argument h=139 plot a horizontal line  at height 139. This is the height result in 4 distinct clustering.***

```{r}
hc.out 
```


```{r}
set.seed(2)
km.out=kmeans(sd.data,4,nstart=20)
km.clusters=km.out$cluster
table(km.clusters,hc.clusters)
```
***We see that the four clusters obtaining hierarchical clustering and K-means clustering are yielding different results. Cluster 2 in K-means clustering is identical to cluster 3 in hierarchical clustering. on the other hand the other clusters are differ.***

***CONCLUSION:*** 

***Hierarchical clustering performs better clustering in NCI60 data set instead of K-means***

